---
title: "Visualize data from [covidtracking.com](http://covidtracking.com)"
author: "Dani Cosme"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: no
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r, include = FALSE}
# load packages
if (!require(tidyverse)) {
  install.packages('tidyverse')
}
```


```{bash, include = FALSE}
# pull data
wget -O daily.csv http://covidtracking.com/api/states/daily.csv
```


```{r, include = FALSE}
# load and tidy data
data = read.csv("daily.csv", stringsAsFactors = FALSE) %>%
  mutate(date = lubridate::ymd(date)) %>%
  gather(var, val, contains("death"), contains("hospital"), contains("positive"), contains("negative")) %>%
  mutate(var = gsub("Increase", "_change", var),
         var = gsub("death", "deaths", var),
         var = gsub("hospitalized", "hospitalizations", var),
         var = gsub("positive", "positive_test_results", var),
         var = gsub("negative", "negative_test_results", var)) %>%
  spread(var, val)
```

# {.tabset}
Here are plots of the current data for the number of deaths, hospitalizations, positive and negative tests for each state over time.

The cumulative plots show the total number in each category, while the log plots show the rate of change in each category. The raw change plots show the daily  increase or decrease. 

```{r}
plot_covid = function(data, y) {
  
  p = data %>%
    ggplot(aes_string("date", y, color = "state")) +
    geom_line() +
    scale_color_manual(values = wesanderson::wes_palette("Darjeeling1", 56, "continuous")) +
    theme_minimal()
  
  if (grepl("change", y)) {
    
    str = gsub("_change", "", y)
    p = p + labs(x = "\ndate", y = sprintf("change in number of %s\n", str))
    
  } else {
    p = p + labs(x = "\ndate", y = sprintf("number of %s\n", gsub("_", " ", y)))
  }
    
  
  plotly::ggplotly(p)
}

plot_covid_log = function(data, y) {
  
  p = data %>%
    mutate(y = log2(!!rlang::parse_quosure(y))) %>%
    ggplot(aes(date, y, color = state)) +
    geom_line() +
    labs(x = "\ndate", y = sprintf("log2 number of %s\n", gsub("_", " ", y))) +
    scale_color_manual(values = wesanderson::wes_palette("Darjeeling1", 56, "continuous")) +
    theme_minimal()
    
  
  plotly::ggplotly(p)
}
```

## deaths {.tabset}
### cumulative
```{r}
plot_covid(data, "deaths")
```

### log
```{r}
plot_covid_log(data, "deaths")
```

### raw change
```{r}
plot_covid(data, "deaths_change")
```

## hospitalizations {.tabset}
### cumulative
```{r}
plot_covid(data, "hospitalizations")
```

### log
```{r}
plot_covid_log(data, "hospitalizations")
```

### raw change
```{r}
plot_covid(data, "hospitalizations_change")
```

## positive test results {.tabset}
### cumulative
```{r}
plot_covid(data, "positive_test_results")
```

### log
```{r}
plot_covid_log(data, "positive_test_results")
```

### raw change
```{r}
plot_covid(data, "positive_test_results_change")
```

## negative test results {.tabset}
### cumulative
```{r}
plot_covid(data, "negative_test_results")
```

### log
```{r}
plot_covid_log(data, "negative_test_results")
```

### raw change
```{r}
plot_covid(data, "negative_test_results_change")
```
